### 2024-10-26 Article/Background_sections/Adhesivity.tex
section on kernel pairs + some proofs

>>>>

\begin{cor}\label{cor:adhesivity_functor_categories}
    Let $\cat A$ be a $\mathcal M$-adhesive category for some $\mathcal M \subseteq \Hom(\cat A)$. Then, for every other category $\cat C$, the functor category $[\cat C, \cat A]$ is $\mathcal M^{\cat C}$-adhesive, where
    \[
        \mathcal{M}^\cat C = \{\eta \in \Hom([\cat C, \cat A]) \mid \eta_C \in \mathcal M \textit{ for each object $C$ of $\cat C$ }\}
    \]
\end{cor}

Since $\Set$ is adhesive, we can conclude what follows.

\begin{cor}\label{cor:presh_adhesive}
    Every category of presheaves is adhesive.
\end{cor}

### 2024-10-26 Article/Background_sections/Basic_Notions.tex
section on kernel pairs + some proofs

@@

\begin{definition}[Product category]
    Given two categories $\cat C, \cat D$, the \emph{product category} $\cat{C\times D}$ has as objects pairs of objects $(A, B)$, where $A \in \Ob(\cat C), B \in \Ob(\cat D)$, and as arrows pairs of morphisms $(f, g)$, where $f$ is an arrow in $\cat C$ and $g$ is an arrow in $\cat D$. Composition and identities are defined pairwise: $(f, g) \circ (h, k) = (f \circ h, g \circ k)$, and $id_{(A, B)} = (id_A, id_B)$.

\end{definition}

>>>>

The next proposition will be useful later.

\begin{prop}\label{prop:prop_epi_mono_Set}
    In $\Set$, for every commutative square as the one below, if $e: X \to Y$ is epi and $m: M \to Z$ is mono, then there exists a unique morphism $h: Y \to M$ making the whole diagram below commutative.
    \[
        \begin{tikzcd}
            X \ar[r, "f"] \ar[d, "e"swap] & M \ar[d, "m"] \\
            Y \ar[ur, dashed, "h"] \ar[r, "g"swap] & Z
        \end{tikzcd}
    \]
\end{prop}

\begin{proof}
    Before we start proving the statement, we have to note that, given a function $t: A \to B$, it is possible to decompose it as a composition of a injective function and a surjective function, considering the function $A \to t(A)$ sending each element onto its image along $t$, and then applying the inclusion $t(A) \to B$, and such functions are unique. Another way to factorize $t$ is via a composition of a surjective function and an injective function. Consider the equivalence relation $\sim$ defined on $A$, where $a \sim a'$ whenever $t(a) = t(a')$. This equivalence relation induces a map $A \to A/_\sim$, which is surjective. The function $A/_\sim \to B$, mapping each equivalence class onto its image along $t$ is then injective, and this factorization in unique too.

    Let now be $f = f_i \circ f_s$ be the decomposition of $f$ with $f_s$ surjective (i.e., epi in $\Set$) and $f_i$ injective (i.e., mono in $\Set$), and $g = g_i \circ g_s$ be the decomposition of $g$ with $g_i$ injective and $g_s$ surjective, having the following situation.
    \[
        \begin{tikzcd}
            X \ar[r, "{f_s}"] \ar[d, "e"] & f(X) \ar[r, "{f_i}"] & M \ar[d, "m"] \\
            Y \ar[r, "{g_i}"swap] \ar[urr, "h"]& Y/_\sim \ar[r, "{g_s}"swap] & Z
        \end{tikzcd}
    \]
    For the diagram above to commute, must be $f_i \circ f_s = h \circ e$ and $g_s \circ g_i = m \circ h$\todo{completare!!!}
\end{proof}

@@

\begin{definition}[Product category]
    Given two categories $\cat C, \cat D$, the \emph{product category} $\cat{C\times D}$ has as objects pairs of objects $(A, B)$, where $A \in \Ob(\cat C), B \in \Ob(\cat D)$, and as arrows pairs of morphisms $(f, g)$, where $f$ is an arrow in $\cat C$ and $g$ is an arrow in $\cat D$. Composition and identities are defined pairwise: $(f, g) \circ (h, k) = (f \circ h, g \circ k)$, and $id_{(A, B)} = (id_A, id_B)$.
\end{definition}

### 2024-10-26 Article/Background_sections/Limit.tex
section on kernel pairs + some proofs

@@

\begin{definition}[Regular Monomorphism]\label{def:reg_mono}
    A monomorphism that is an equalizer for a pair of arrows is said \emph{regular monomorphism}. The class of all regular monomorphisms of a category $\cat C$ is denoted $\Reg(\cat C)$.
\end{definition}

@@

@@

>>>>

The dual of the proposition above states that a coequalizer is an epimmorphism.

@@

\begin{definition}[Regular Monomorphism, Regular Epimorphism]\label{def:reg_mono}
    A monomorphism that is an equalizer for a pair of arrows is said \emph{regular monomorphism}. The class of all regular monomorphisms of a category $\cat C$ is denoted $\Reg(\cat C)$. An epimorphism that is a coequalizer for a pair of arrows is said \emph{regular epimorphism}.
\end{definition}

@@

\begin{cor}\label{cor:cube} 
    Let $\cat{C}$ be a category and suppose that  the solid part of the following cube is given
	\[\begin{tikzcd}[row sep=13 pt, column sep=13 pt]
    	& {Y'} && {Z'} \\
    	{B'} && {C'} \\
    	& Y && Z \\
    	B && C
    	\arrow["{g'}", from=1-2, to=1-4]
    	\arrow["{q'}"', dashed, from=1-2, to=2-1]
    	\arrow["y"'{pos=0.8}, from=1-2, to=3-2]
    	\arrow["{r'}", from=1-4, to=2-3]
    	\arrow["z", from=1-4, to=3-4]
    	\arrow["{k'}"{pos=0.8}, from=2-1, to=2-3, crossing over]
    	\arrow["b"', from=2-1, to=4-1]
    	\arrow["c"{pos=0.7}, from=2-3, to=4-3]
    	\arrow["g"{pos=0.3}, from=3-2, to=3-4]
    	\arrow["q"'{pos=0.2}, from=3-2, to=4-1]
    	\arrow["r", from=3-4, to=4-3]
    	\arrow["k"', from=4-1, to=4-3]
        \end{tikzcd}\]
	If the front face is a pullback then there is a unique $q'\colon Y'\to B'$ filling the diagram. If, moreover, the other two vertical faces are also pullbacks, then the following square is a pullback too.
	\[\begin{tikzcd}
	    Y' \ar[r, "{q'}"] \ar[d, "y"swap] & B' \ar[d,"b"] \\
            Y \ar[r, "q"swap] & B
	\end{tikzcd}\]
\end{cor}
\begin{proof}
	Let us compute:
	\begin{align*}
            c\circ r'\circ g'   &=  r\circ z \circ g'   \\
                                &=  r\circ g \circ y    \\
                                &=k\circ q \circ y
	\end{align*}
	Since the front face is a pullback, this guarantees the existence of $q'$.  The second half of the thesis follows applying \Cref{lemma:pullback_lemma} to the following rectangle.
	\[\begin{tikzcd}[row sep=15 pt, column sep=15 pt]
	{Y'} && {B'} && {C'} \\
	\\
	Y && B && C
	\arrow["{q'}"', from=1-1, to=1-3]
	\arrow["{r' \circ g'}", bend left = 30, from=1-1, to=1-5]
	\arrow["y"', from=1-1, to=3-1]
	\arrow["{k'}"', from=1-3, to=1-5]
	\arrow["b", from=1-3, to=3-3]
	\arrow["c", from=1-5, to=3-5]
	\arrow["q", from=3-1, to=3-3]
	\arrow["{r \circ g}"', bend right=30, from=3-1, to=3-5]
	\arrow["k", from=3-3, to=3-5]
\end{tikzcd}\]
\end{proof}

@@

\subsection{Kernel pairs}

\begin{definition}[Kernel Pair]
    A \emph{kernel pair} for an arrow $f: A \to B$ is an object $K_f$ together with two arrows $\pi_1^f, \pi_2^f : K_f \to A$, denoted as $(K_f, \pi_1^f, \pi_2^f)$, such that the following square is a pullback.
    \[
        \begin{tikzcd}
            K_f \ar[r, "{\pi_1^f}"] \ar[d, "{\pi_2^f}" swap] & A \arrow[d, "f"] \\
            A \ar[r, "f"swap] & B
        \end{tikzcd}
    \]
\end{definition}

\begin{remark}
	If a category $\cat{C}$ has pullbacks then every arrow has a kernel pair.
\end{remark}

\begin{prop}\label{prop:kermono}
	An arrow $m\colon M\to X$ is mono if and only if $(M, id_M, id_M)$ is a kernel pair for it.
\end{prop}

\begin{proof}
    To prove the ``if'' part of the statement, let $f, g: A \to M$ be such that $m\circ f = m\circ g$, and consider the following situation.
    \[
        \begin{tikzcd}
        A \arrow[drr, bend left=20, "f"] \arrow[ddr, bend right=20,"g" swap] \arrow[dr, dashed, "u"] & & \\
        & M  \arrow[d, "{id_M}" swap] \arrow[r, "{id_M}"] & M \arrow[d, "m"] \\
        & M  \arrow[r, "m" swap] & X
        \end{tikzcd}
    \]
    For the universal property of pullbacks, we have that $$f  =  id_M \circ u =  g$$
    Hence, $m$ is mono.

    Conversely, if $m$ is mono, then, we have that
    \begin{align*}
        m \circ f = m \circ g   &\Rightarrow    f = g \\
                                &\Rightarrow    f \circ id_M = g\circ id_M
    \end{align*}
    Hence, $f$ is the unique arrow that makes the commutative square illustrated above a pushout.
\end{proof}

\begin{cor}\label{cor:kermono}
	Let $(K_f, \pi^f_i), \pi^f_2)$ be a kernel pair for $f\colon X\to Y$. Then for every mono $m\colon Y\to Z$, $(K_f, \pi^f_1, \pi^f_2)$ is a kernel pair also for $m\circ f$.
\end{cor}
\begin{proof}
    It is enough to see that, by \Cref{lemma:pullback_lemma} and \Cref{prop:kermono} the outer boundary of the following square is a pullback.
        \[\begin{tikzcd}[row sep=13 pt, column sep=13 pt]
    	{K_f} && X && X \\
    	\\
    	X && Y && Y \\
    	\\
    	X && Y && Z
    	\arrow["{\pi_f^2}", from=1-1, to=1-3]
    	\arrow["{\pi_f^2}"', from=1-1, to=3-1]
    	\arrow["{id_X}", from=1-3, to=1-5]
    	\arrow["f", from=1-3, to=3-3]
    	\arrow["f", from=1-5, to=3-5]
    	\arrow["f", from=3-1, to=3-3]
    	\arrow["{id_X}"', from=3-1, to=5-1]
    	\arrow["{id_Y}", from=3-3, to=3-5]
    	\arrow["{id_Y}", from=3-3, to=5-3]
    	\arrow["m", from=3-5, to=5-5]
    	\arrow["f"', from=5-1, to=5-3]
    	\arrow["m"', from=5-3, to=5-5]
    \end{tikzcd}\]
\end{proof}

\begin{lemma}
    Suppose the following situation, and that $f: X \to Y$ and $g: Z \to W$ have kernel pairs.
    \[
        \begin{tikzcd}
            X \ar[r, "h"] \ar[d, "f"swap] & Z \ar[d, "g"] \\
            Y \ar[r, "t"swap] & W
        \end{tikzcd}
    \]

    Then, there exists a unique arrow $k_h: K_f \to K_g$ making the squares below commute.
    \[
        \begin{tikzcd}[row sep = 25 pt, column sep= 25 pt]
            K_f \ar[r, dashed, "{k_h}"] \ar[d, "{\pi_f^1}"swap] & K_g \ar[d, "{\pi_g^1}"] \\
            X \ar[r, "h"swap] & Z 
        \end{tikzcd}
        \qquad
        \begin{tikzcd}[row sep = 25 pt, column sep= 25 pt]
            K_f \ar[r, dashed, "{k_h}"] \ar[d, "{\pi_f^2}"swap] & K_g \ar[d, "{\pi_g^2}"] \\
            X \ar[r, "h"swap] & Z 
        \end{tikzcd}
    \]

    Moreover, if the beginning square is a pullback, then also the preceding ones are so.
\end{lemma}

\begin{proof}
    Computing, we have
    \begin{align*}
        g \circ h \circ \pi_f^1 &=  t \circ f \circ \pi_f^1     \\
                                &=  t \circ f \circ \pi_f^2     \\
                                &=  g \circ h \circ \pi_f^2
    \end{align*}
    By the universal property of $K_g$ as the pullback of $g$ along itself, such $k_h$ exists and it is unique.

    To prove the second half of the thesis, let us consider the two rectangles below, which, by \Cref{lemma:pullback_lemma} are pullbacks.
    \[\begin{tikzcd}[row sep= 25 pt, column sep= 25 pt]
	{K_f} & X & Z \\
	X & Y & W
	\arrow["{\pi_f^1}", from=1-1, to=1-2]
	\arrow["{\pi_f^2}"', from=1-1, to=2-1]
	\arrow["h", from=1-2, to=1-3]
	\arrow["f", from=1-2, to=2-2]
	\arrow["g", from=1-3, to=2-3]
	\arrow["f"', from=2-1, to=2-2]
	\arrow["t"', from=2-2, to=2-3]
    \end{tikzcd}
    \qquad
    \begin{tikzcd}[row sep = 25 pt, column sep=25 pt]
	{K_f} & X & Z \\
	X & Y & W
	\arrow["{\pi_f^2}", from=1-1, to=1-2]
	\arrow["{\pi_f^1}"', from=1-1, to=2-1]
	\arrow["h", from=1-2, to=1-3]
	\arrow["f", from=1-2, to=2-2]
	\arrow["g", from=1-3, to=2-3]
	\arrow["f"', from=2-1, to=2-2]
	\arrow["t"', from=2-2, to=2-3]
    \end{tikzcd}
    \]
    But then the following ones are pullbacks too.
    \[\begin{tikzcd}[row sep= 25 pt, column sep= 25 pt]
	{K_f} & {K_g} & Z \\
	X & Y & W
	\arrow["{k_h}"', from=1-1, to=1-2]
	\arrow["{\pi_f^1}"', from=1-1, to=2-1]
	\arrow["{\pi_g^2}"', from=1-2, to=1-3]
	\arrow["{\pi_g^1}", from=1-2, to=2-2]
	\arrow["g", from=1-3, to=2-3]
	\arrow["h", from=2-1, to=2-2]
	\arrow["g", from=2-2, to=2-3]
        \arrow["{h\circ \pi_f^2}", from=1-1, to=1-3, bend left = 30]
        \arrow["{t \circ f}"', from=2-1, to=2-3, bend right = 30]
    \end{tikzcd}
    \qquad
    \begin{tikzcd}[row sep = 25 pt, column sep=25 pt]
	{K_f} & {K_g} & Z \\
	X & Y & W
	\arrow["{k_h}"', from=1-1, to=1-2]
	\arrow["{\pi_f^2}"', from=1-1, to=2-1]
	\arrow["{\pi_g^1}"', from=1-2, to=1-3]
	\arrow["{\pi_g^2}", from=1-2, to=2-2]
	\arrow["g", from=1-3, to=2-3]
	\arrow["h", from=2-1, to=2-2]
	\arrow["g", from=2-2, to=2-3]
        \arrow["{h\circ \pi_f^1}", from=1-1, to=1-3, bend left = 30]
        \arrow["{t \circ f}"', from=2-1, to=2-3, bend right = 30]
    \end{tikzcd}
    \]

    The thesis follows again by \Cref{lemma:pullback_lemma}.
\end{proof}

\begin{cor}
    Let $\cat C$ be a category with pullbacks and $\phi : D \to D'$ be a natural transformation between two functors $D, D': \cat{I \to C}$. If $\phi_i$ is a regular epi for every $i$, then $\phi$ is a regular epi.
\end{cor}

\begin{proof}
    TODO
\end{proof}
